---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cardid

<!-- badges: start -->
<!-- badges: end -->

The goal of cardid is to allow you to use the information of basketball and 
    baseball sports trading cards in order to find the (1) player and (2) year
    that each card refers to. In addition, it allows the creation of a card 
    identifier.

## Installation

You can install the released version of cardid from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("cardid")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jk2049/cardid")
```
## Example

Below you can find an example of the functionalities offered. Please keep in 
mind that this is an early version. As a result, the results of this code have 
only been tested when running the functions in the same order as in the below 
example:

```{r example}
library(cardid)
## basic example code
```

The package offers three datasets. The first is a dataset with information
about all the player who have played in the MLB:

```{r dataset1}
head(baseball.players)
```

The second is a dataset with information about all the player who have 
played in the NBA:
```{r dataset2}
head(basketball.players)
```

The third is a subset of a dataset that contains information about sports 
trading cards that were listed on eBay:
```{r dataset3}
head(baseball.players)
head(basketball.players)
head(dt.example)

dt.example.titles <- dt.example[, list(sports, card_title)]
dt.example.basketball.titles <- dt.example.titles[sports == "Basketball", ]
dt.example.baseball.titles <- dt.example.titles[sports == "Baseball", ]
dt.example.years <- dt.example[, list(sports, card_title, product_section_year)]
rm(dt.example, dt.example.titles)
```

You'll still need to render `README.Rmd` regularly, to keep `README.md` up-to-date. `devtools::build_readme()` is handy for this. You could also use GitHub Actions to re-render `README.Rmd` every time you push. An example workflow can be found here: <https://github.com/r-lib/actions/tree/master/examples>.

You can also embed plots, for example:

```{r pressure, echo = FALSE}
plot(pressure)
```

In that case, don't forget to commit and push the resulting figure files, so they display on GitHub and CRAN.
