---
output: github_document
---

<!-- README.md is generated from README.Rmd. Please edit that file -->

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>",
  fig.path = "man/figures/README-",
  out.width = "100%"
)
```

# cardid

<!-- badges: start -->
<!-- badges: end -->

The goal of cardid is to allow you to use the information of basketball and 
    baseball sports trading cards in order to find the (1) player and (2) year
    that each card refers to. In addition, it allows the creation of a card 
    identifier.

## Installation

You can install the released version of cardid from [CRAN](https://CRAN.R-project.org) with:

``` r
install.packages("cardid")
```

And the development version from [GitHub](https://github.com/) with:

``` r
# install.packages("devtools")
devtools::install_github("jk2049/cardid")
```
## Example

Below you can find an example of the functionalities offered. Please keep in 
mind that this is an early version. As a result, the results of this code have 
only been tested when running the functions in the same order as in the below 
example:

```{r example}
library(cardid)
library(dplyr)
library(stringr)
library(data.table)
library(plyr)
```

The package offers three datasets. The first is a dataset with information
about all the player who have played in the MLB:

```{r dataset1}
head(dt.baseball.players)
```

The second is a dataset with information about all the player who have 
played in the NBA:
```{r dataset2}
head(dt.basketball.players)
```

The third is a subset of a dataset that contains information about sports 
trading cards that were listed on eBay:
```{r dataset3}
head(dt.example)

dt.example.titles <- dt.example[, list(sports, card_title)]
dt.example.basketball.titles <- dt.example.titles[sports == "Basketball", ]
dt.example.baseball.titles <- dt.example.titles[sports == "Baseball", ]
dt.example.years <- dt.example[, list(sports, card_title, product_section_year)]
```

### Player Name Pre-Processing

First, we use the functions ''baseball.preprocess'' and ''baseball.preprocess''
in order to manipulate the player names of the MLB and the NBA players 
respectively. Standard NLP actions are taken (i.e., remove stopwords etc.) that
allow us to build a dictionary of player names. This is important as the same 
player names (e.g., shaquille o'neal) can be written in different ways in the
card titles (e.g., shaquille o'neal, shaquille o neal, shaquille oneal etc.).

```{r playernamepreprocessing}
baseball.players <- baseball.preprocess(dt.baseball.players)
basketball.players <- basketball.preprocess(dt.basketball.players)
```

### Card Title Pre-Processing

Second, we use the function ''data.preprocess'' on the card title columns in 
order to manipulate the card titles (i.e., remove stopwords etc.)

```{r cardtitlepreprocessing}
dt.example.basketball.titles$card_title <- data.preprocess(dt.example.basketball.titles$card_title)
dt.example.baseball.titles$card_title <- data.preprocess(dt.example.baseball.titles$card_title)
dt.example.years$card_title_2 <- data.preprocess(dt.example.years$card_title)
```

### Get Card Player

Then, we use the function ''get.player'' to find the player(s) that each card
title refers to. 

```{r echo=FALSE, results='hide',message=FALSE}
get.player <- function(dt.input, card.title.col.name = "card_title", sport){

  card.title.index <- which(names(dt.input) == card.title.col.name)

  if(sport == "basketball"){
    players.in.title <- basketball.players[str_detect(dt.input[card.title.index], basketball.players$player),
                                           "original_player"]
  }else if(sport == "baseball"){
    players.in.title <- baseball.players[str_detect(dt.input[card.title.index], baseball.players$player),
                                         "original_player"]
  }else{
    stop(paste("Sport ", sport, " is not supported."))
  }
  n.players <- length(players.in.title)
  players.in.title <- paste(players.in.title, sep = ",")

  if(length(players.in.title) == 0) players.in.title <- NA

  dt.output <- c(dt.input, "n_players" = n.players,
                 players.in.title)
  dt.output <- as.data.frame(t(dt.output))

  return(dt.output)
}

```

```{r getplayer}
dt.example.basketball.titles <- rbindlist(
  apply(dt.example.basketball.titles, 1, get.player, 
        card.title.col.name = "card_title", sport = "basketball"), fill=TRUE)
dt.example.baseball.titles <- rbindlist(
  apply(dt.example.baseball.titles, 1, get.player, 
        card.title.col.name = "card_title", sport = "baseball"), fill=TRUE)
```

### Get Card Year

The next step is to identify the year that each card refers to. We can identify
the card year with at most two ways: One is by analyzing the card title through
the function ''get.year.from.title'', while the other is to analyze the 
information contained in the product section field of the listed cards through 
the function ''get.year.from.product.section''. The product section field may 
or may not have been filled in by the seller. 

Once we are done, we use ''get.product.year'' to construct the card year column.

```{r getyear}
# Year from the card title
dt.example.years <- suppressWarnings(rbindlist(apply(dt.example.years, 1, get.year.from.title,
                                                     card.title.col.name = "card_title"), fill=TRUE))

# Year from the product section information
dt.example.years <- rbindlist(apply(dt.example.years, 1, get.year.from.product.section,
                                    product.section.year.col.name = "product_section_year"), fill=TRUE)

# Combine the two
dt.example.years <- get.product.year(dt.example.years)
```

### Get Product ID 

Now that we have the identified player and year of every card, we can construct
the product identifier. If there is no identified player or no identified year
for a card, then that we insert a missing value (i.e., <NA>) in that card's 
product id column.

```{r getproductid}
dt.example.names <- rbind.fill(dt.example.baseball.titles, dt.example.basketball.titles)
rm(dt.example.baseball.titles, dt.example.basketball.titles)

dt.example.names <- as.data.table(dt.example.names)
dt.example.names <- dt.example.names[!duplicated(dt.example.names)]

dt.example <- merge(dt.example.years, dt.example.names,
                    by.x = c("sports", "card_title_2"),
                    by.y = c("sports", "card_title"), all.x = TRUE)

dt.example[, product_id := paste0(V4, "-", product_id_year)]
# If any component is NA then replace the ID with NA
dt.example[is.na(V4), product_id := NA]
dt.example[is.na(product_id_year), product_id := NA]
```

### Get Valid Product ID

Although we have constructed a product identifier, there are still some
pending issues that can only be discovered upon closer inspection of the data:
  <ol>
  <li> For some entries the previous functions have identified identical player
names that have been inserted in separate columns. By using 
''remove.duplicate.players'' we can remove the duplicates and only keeps
the unique names</li>
  <li> There's an issue with some names that contain "jr.". This issue is only 
present for the baseball cards. We address this issue using ''fix.names''. </li>
<li> In some cases the identified player names can refer to players who were 
active during different periods in time (e.g., Bob Smith for baseball). We 
address this issue by using ''player.tie.break''. This function utilizes the 
identified card year and pinpoints which specific player each card refers to. </li>
<li> We have constructed a product id for any cards for which we have identified
at least one player and a year. The product id however should only be valid 
for those cards for which exactly one player has been identified. The function
''flag.id.valid'' flags these observations. </li>
</ol>


```{r echo=FALSE, results='hide',message=FALSE}
remove.duplicate.players <- function(dt.input, sport){

  # Subset data to specific sport
  dt.input$sports <- tolower(dt.input$sports)
  dt.input <- dt.input[sports == sport, ]

  if(sport == "basketball"){
    all.players <- basketball.players$original_player
    ## Keep duplicate entries
    all.players.duplicates <- all.players[duplicated(all.players)]
  }else if(sport == "baseball"){
    all.players <- baseball.players$original_player
    ## Keep duplicate entries
    all.players.duplicates <- all.players[duplicated(all.players)]
    all.players.duplicates <- c(all.players.duplicates,
                                all.players[grepl("jr\\.", all.players)])
  }else{
    stop(paste("Sport ", sport, " is not supported."))
  }
  rm(all.players)

  i.limit <- as.numeric(max(dt.input$n_players)) + 3

  dt.input$flag <- 0
  dt.input$sports <- tolower(dt.input$sports)


  for(i in 4:i.limit){
    col.player.name <- paste0("V", i)
    dt.input[get(col.player.name) %in% all.players.duplicates, flag := 1]
    rm(col.player.name)
  }
  indices <- which(dt.input$flag == 1)

  # Fix columns
  # Some V columns have the same players
  for(i in 1:length(indices)){

    print(paste0("Loop 1 - ", sport, " ", i, " out of ", length(indices)))

    card.info <- dt.input[indices[i], ]

    # Get all the players and insert them into a vector
    players.in.title <- c()
    for(j in 4:i.limit){
      col.player.name <- paste0("V", j)
      players.in.title <- append(players.in.title, card.info[, get(col.player.name)])
      rm(col.player.name)
    }
    # Remove the NAs
    players.in.title <- players.in.title[!is.na(players.in.title)]
    # Get the number of duplicate players
    n.duplicated.player.in.title <- sum(duplicated(players.in.title))
    # Get the duplicate players
    duplicated.player.in.title <- players.in.title[duplicated(players.in.title)]

    # For every duplicate player in the vector
    j <- 1
    # Insert all the players into a new vector
    insert.players.in.title <- players.in.title
    while(j <= length(duplicated.player.in.title)){

      # Find the index of all the entries of that player
      duplicated.player.in.title.index <- which(players.in.title == duplicated.player.in.title[j])
      # Remove the first index (we want to keep the first - i.e. one - entry)
      duplicated.player.in.title.index <- duplicated.player.in.title.index[-1]
      # Replace the duplicate indices with NA
      insert.players.in.title[duplicated.player.in.title.index] <- NA

      rm(duplicated.player.in.title.index)
      j <- j + 1
    }
    # Remove the NAs
    insert.players.in.title <- insert.players.in.title[!is.na(insert.players.in.title)]

    # Change n_players
    dt.input[indices[i], n_players := as.numeric(length(insert.players.in.title))]

    # Lengthen the new vector so that it has i.limit elements (corresponding to V4-Vi.limit)
    insert.players.in.title <- c(insert.players.in.title,
                                 rep(NA, i.limit-length(insert.players.in.title)))
    # Insert the new elements into the current columns
    for(z in 1:i.limit){
      col.player.name <- paste0("V", z+3)
      dt.input[indices[i], (col.player.name) := insert.players.in.title[z]]
      rm(col.player.name)
    }

    rm(card.info, players.in.title,
       n.duplicated.player.in.title, duplicated.player.in.title,
       j, z, insert.players.in.title)
  }
  return(dt.input)
}

fix.names <- function(dt.input, sport){
  if(sport == "baseball"){
    dt.input <- dt.input[sports == sport, ]

    i.limit <- as.numeric(max(dt.input$n_players)) + 3
    dt.input$flag <- 0
    for(i in 4:i.limit){
      col.player.name <- paste0("V", i)
      dt.input[grepl(" jr\\.", get(col.player.name)), flag := 1]
      rm(col.player.name)
    }
    indices <- which(dt.input$flag == 1)

    for(i in 1:length(indices)){

      print(paste0("Loop 2 - sport ", sport, " ", i, " out of ", length(indices)))

      card.info <- dt.input[indices[i], ]

      # Get all the players and insert them into a vector
      players.in.title <- c()
      for(j in 4:i.limit){
        col.player.name <- paste0("V", j)
        players.in.title <- append(players.in.title, card.info[, get(col.player.name)])
        rm(col.player.name)
      }
      # Remove the jr. part
      players.in.title.jr.removed <- gsub(" jr\\.", "", players.in.title)
      # Remove the NAs
      players.in.title <- players.in.title[!is.na(players.in.title)]
      players.in.title.jr.removed <- players.in.title.jr.removed[!is.na(players.in.title.jr.removed)]
      # Get the number of duplicate players
      n.duplicated.player.in.title <- sum(duplicated(players.in.title.jr.removed))
      # Get the duplicate players
      duplicated.player.in.title <- players.in.title.jr.removed[duplicated(players.in.title.jr.removed)]

      # For every duplicate player in the vector
      j <- 1
      # Insert all the players into a new vector
      insert.players.in.title <- players.in.title
      while(j <= length(duplicated.player.in.title)){

        # Find the index of all the entries of that player
        duplicated.player.in.title.index <- which(grepl(duplicated.player.in.title[j], players.in.title))
        # Remove the first index (we want to keep the first - i.e. one - entry)
        duplicated.player.in.title.index <- duplicated.player.in.title.index[-2]
        # Replace the duplicate indices with NA
        insert.players.in.title[duplicated.player.in.title.index] <- NA

        rm(duplicated.player.in.title.index)
        j <- j + 1
      }
      # Remove the NAs
      insert.players.in.title <- insert.players.in.title[!is.na(insert.players.in.title)]

      # Change n_players
      dt.input[indices[i], n_players := as.numeric(length(insert.players.in.title))]

      # Lengthen the new vector so that it has 8 elements (corresponding to V4-V11)
      insert.players.in.title <- c(insert.players.in.title,
                                   rep(NA, 8-length(insert.players.in.title)))
      # Insert the new elements into the current columns
      for(z in 1:i.limit){
        col.player.name <- paste0("V", z+3)
        dt.input[indices[i], (col.player.name) := insert.players.in.title[z]]
        rm(col.player.name)
      }

      rm(card.info, players.in.title,
         n.duplicated.player.in.title, duplicated.player.in.title,
         j, insert.players.in.title, z)
    }
    dt.input[, flag := NULL]
    return(dt.input)

  }else{
    stop(paste("This function is not valid for sport ", sport, ". It's only valid for baseball."))
  }
}

player.tie.break <- function(dt.input, sport){

  if(sport == "basketball"){
    all.players <- dt.basketball.players$player
  }else if(sport == "baseball"){
    all.players <- dt.baseball.players$player
  }else{
    stop(paste("Sport ", sport, " is not supported."))
  }
  all.players.duplicates <- all.players[duplicated(all.players)]
  if(sport == "basketball"){
    all.players <- dt.basketball.players
  }else if(sport == "baseball"){
    # Reconstruct flag
    dt.input$flag <- 0
    i.limit <- as.numeric(max(dt.input$n_players)) + 3
    for(i in 4:i.limit){
      col.player.name <- paste0("V", i)
      dt.input[get(col.player.name) %in% all.players.duplicates, flag := 1]
      rm(col.player.name)
    }
    rm(i, i.limit)

    all.players <- dt.baseball.players
  }

  indices <- which(dt.input$flag == 1)

  for(i in 1:length(indices)){

    print(paste0("Loop 3 - ", sport, " player ", i, " out of ", length(indices)))

    card.info <- dt.input[indices[i], ]
    card.title <- card.info[, card_title_2]

    # Get all the players and insert them into a vector
    players.in.title <- c()
    col.names <- c()
    i.limit <- as.numeric(max(dt.input$n_players)) + 3
    for(j in 4:i.limit){
      col.player.name <- paste0("V", j)
      col.names <- append(col.names, col.player.name)
      players.in.title <- append(players.in.title, card.info[, get(col.player.name)])
      rm(col.player.name)
    }

    # Keep only the names that are in the names of the duplicated vector
    players.in.title <- players.in.title[players.in.title %in% all.players.duplicates]

    players.in.title.count <- 1
    while(players.in.title.count <= length(players.in.title)){
      # Get active years for every player
      player.active.year <- all.players[player == players.in.title[players.in.title.count], list(player, from, to)]
      # Compare with card title year
      player.active.year[, potential := ((as.numeric(from)-1) <= as.numeric(card.info[, product_id_year]) &
                                           as.numeric(to) >= as.numeric(card.info[, product_id_year]))]

      player.count <- sum(player.active.year$potential == TRUE)
      player.index <- which(player.active.year$potential == TRUE)
      if(!is.na(player.count) & player.count == 1){
        # One player fits
        player.name <- player.active.year[potential == TRUE, player]
        new.player.name <- paste0(player.name, "-", player.index)
      }else if(!is.na(player.count) & player.count == 0){
        # No players fit
        player.name <- player.active.year[1, player]
        new.player.name <- paste0(player.name, "-no-fit")
      }else if(!is.na(player.count) & player.count > 1){
        # All players fit
        player.name <- player.active.year[1, player]
        new.player.name <- paste0(player.name, "-all-fit")
      }else if(is.na(player.count)){
        # All players fit
        player.name <- player.active.year[1, player]
        new.player.name <- paste0(player.name, "-NA")
      }
      rm(player.active.year, player.count, player.index)

      pot.cols <- card.info[, col.names, with = FALSE]
      col.ind <- which(pot.cols == player.name)
      rm(pot.cols, player.name)

      col.player.name <- paste0("V", col.ind+3)
      dt.input[indices[i], (col.player.name) := new.player.name]
      rm(new.player.name, col.ind, col.player.name)

      players.in.title.count <- players.in.title.count + 1
    }
    rm(card.info, card.title, players.in.title, players.in.title.count)
    rm(col.names)
  }
  dt.input$flag <- NULL
  return(dt.input)
}

flag.id.valid <- function(dt.input){
  # Flag valid
  dt.input[, valid := 0]
  dt.input[n_players == 1 &
             !grepl("-no-fit", V4) & !grepl("-all-fit", V4) & !grepl("-NA", V4) &
             !grepl("-no-fit", V5) & !grepl("-all-fit", V5) & !grepl("-NA", V5) &
             !grepl("-no-fit", V6) & !grepl("-all-fit", V6) & !grepl("-NA", V6) &
             !grepl("-no-fit", V7) & !grepl("-all-fit", V7) & !grepl("-NA", V7) &
             !grepl("-no-fit", V8) & !grepl("-all-fit", V8) & !grepl("-NA", V8) &
             !grepl("-no-fit", V9) & !grepl("-all-fit", V9) & !grepl("-NA", V9), valid := 1]
  dt.input[is.na(product_id_year), valid := 0]

  return(dt.input)
}

```

```{r getvalidproductid}
dt.example.5.1.1 <- remove.duplicate.players(dt.example, "basketball")
dt.example.5.1.2 <- remove.duplicate.players(dt.example, "baseball")

dt.example.5.2.2 <- fix.names(dt.example.5.1.2, "baseball")

dt.example.5.3.1 <- player.tie.break(dt.example.5.1.1, "basketball")
dt.example.5.3.2 <- player.tie.break(dt.example.5.2.2, "baseball")

dt.example.5 <- rbind.fill(dt.example.5.3.1, dt.example.5.3.2)
dt.example.5 <- as.data.table(dt.example.5)

dt.example.valid <- flag.id.valid(dt.example.5)
```

### Drop Columns

Using the function ''drop.useless.columns'' we can drop some columns that are
no longer needed. 

```{r echo=FALSE, results='hide',message=FALSE}
drop.useless.columns <- function(dt.input){

  # Columns that are not needed any longer are dropped
  dt.input$four_digit_year <- NULL
  dt.input$four_digit_year_count <- NULL
  dt.input$two_digit_year <- NULL
  dt.input$two_digit_year_count <- NULL
  dt.input$two_to_four_digit_year <- NULL
  dt.input$four_digit_product_section_year <- NULL
  dt.input$four_digit_product_section_year_count <- NULL
  dt.input$two_digit_product_section_year <- NULL
  dt.input$two_digit_product_section_year_count <- NULL
  dt.input$two_to_four_digit_product_section_year <- NULL

  # Also drop the V columns that are empty
  for(i in 4:11){
    col.player.name <- paste0("V", i)
    if(col.player.name %in% colnames(dt.input)){
      if(sum(is.na(dt.input[, get(col.player.name)])) == nrow(dt.input)){
        dt.input[, (col.player.name) := NULL]
      }
      rm(col.player.name)
    }
  }
  return(dt.input)
}

```

```{r dropcolumns}
colnames(dt.example.valid)
dt.example.valid <- drop.useless.columns(dt.example.valid)
colnames(dt.example.valid)
```

### Get Additional Player Information

Our initial sports player datasets contain information that we have not 
included. Using the function ''add.player.info'' we add this information to our
dataset.

```{r echo=FALSE, results='hide',message=FALSE}
add.player.info <- function(dt.input){

  # dt.input <- dt.example.valid[1347, ]

  baseball.players.keep <- baseball.players
  basketball.players.keep <- basketball.players

  baseball.players$player <- baseball.players$original_player
  baseball.players$original_player <- NULL
  baseball.players <- baseball.players[!duplicated(baseball.players), ]
  basketball.players$player <- basketball.players$original_player
  basketball.players$original_player <- NULL
  basketball.players <- basketball.players[!duplicated(basketball.players), ]
  dt.baseball.players <- as.data.table(baseball.players)
  dt.basketball.players <- as.data.table(basketball.players)
  dt.basketball.players[, player := tolower(player)]
  dt.baseball.players[, player := tolower(player)]

  # # Find how many times each name appears
  dt.basketball.players[, n := .N, by = player]
  dt.baseball.players[, n := .N, by = player]

  # # Create an index for the names
  dt.basketball.players[, index := 1:.N, by = player]
  dt.baseball.players[, index := 1:.N, by = player]
  # dt.baseball.players[n == 5, ]

  # # For those names that appear more than once change name to name-index
  dt.basketball.players[, player_name := player]
  dt.baseball.players[, player_name := player]
  dt.basketball.players$A <- 1
  dt.basketball.players$A <- NULL
  dt.baseball.players$A <- 1
  dt.baseball.players$A <- NULL

  dt.basketball.players[n != 1, player_name := paste0(player_name, "-", index)]
  dt.baseball.players[n != 1, player_name := paste0(player_name, "-", index)]

  dt.basketball.players$sports <- "basketball"
  dt.baseball.players$sports <- "baseball"

  dt.players <- rbind(dt.basketball.players[, list(sports, player_name, from, to, active, hall_of_famer)],
                      dt.baseball.players[, list(sports, player_name, from, to, active, hall_of_famer)])

  col.names <- colnames(dt.input)
  col.names <- c(col.names, colnames(dt.players))
  col.names <- unique(col.names)
  col.names <- col.names[col.names != "player_name"]

  # Use this new name to add HoF to dt.cards
  dt.input <- merge(dt.input, dt.players,
                    by.x = c("sports", "V4"),
                    by.y = c("sports", "player_name"),
                    all.x = TRUE)
  setcolorder(dt.input, col.names)

  return(dt.input)
}

```

```{r getadditionalinfo}
dt.example.valid.hof <- add.player.info(dt.example.valid)
head(dt.example.valid.hof)
```

### Data Inspection 

Let's inspect the important columns of our produced dataset. The first column 
(''card_title'') shows the original card title. The second (''n_players'') shows
the number of players that were identified using our dictionary, while the third 
(''V4'') shows the first player name that was identified by our dictionary. If
there is more than one identified player, then each following name is inserted
into columns ''V5'', ''V6'', etc. (these columns are not displayed here to save
space). In the fourth column (''product_id_year'') we can find the identified 
card year, while in the fifth column (''product_id'') we see the product id. In 
the sixth column we can find whether the product id is valid (''valid'' equals 
1). A product id is valid only if (1) a year has been identified for this card
and (2) exactly one player name has been identified for this same card. In the 
columns that follow we can find the date that the identified player started 
(''from'') and stopped (''to'') playing in the MLB/NBA, as well as whether the
identified player is currently active in the MLB/NBA (''active'' equals 1) and 
whether the player has been inducted to the hall of fame (''hall of famer'' 
equals 1).

```{r datainspection}
dt.example.valid.hof <- dt.example.valid.hof[, list(card_title, n_players, V4, 
                                 product_id_year, product_id, 
                                 valid, from, to, active, hall_of_famer)]
head(dt.example.valid.hof)

```

table(dt.example.valid.hof$valid)
head(dt.example.valid.hof[valid == 1, ])
head(dt.example.valid.hof[valid == 0, ])
